Functional Analysis of CGM Data

### 2022-12-02 StatNet/cgm_fda.qmd
Update cgm_fda.qmd

@@

```{r}
basisobj = create.fourier.basis(rangeval=c(0, 1))
plot(basisobj)
```

```{r cache=TRUE}
fpca = rfr_fpca(Y = "Glucose", data = fda_df)
fda_df %>%
  modelr::add_predictions(fpca, var = "glucose_fits") %>%
  ggplot(aes(y = glucose_fits)) +
  geom_spaghetti() +
  geom_spaghetti(aes(y = Glucose), color = "blue", alpha = .2)
```

```{r}
ys = apply(sleep,1,function(r){
  # Pull info
  id = as.character(r["Subject.ID"])
  start = ymd_hms(as.character(r["Night.Start"]))
  end = ymd_hms(as.character(r["Night.End"]))
  # Find CGM file
  df = read.csv(cgms[id])
  id = paste(as.character(r["Subject.ID"]),as.character(r["Day"]),sep = "_")
  df$timestamp = round_date(ymd_hms(df$timestamp),"5 minutes")
  # Get the time between sensors
  interval = table(round(diff(df$timestamp)/60))
  interval = names(interval)[which.max(interval)]
  # Put glucose values on time grid
  times = round_date(seq(start,end,by = paste(interval,"min")),"5 minutes")
  glucs = df$sensorglucose[match(times,df$timestamp)]
  y = data.frame(".obs"=rep(id,length(glucs)),
                 ".index"=seq(0,1,length.out=length(glucs)),
                 ".value" = glucs)
  return(y)
})
# Find those where all glucose values are missing
# keep = sapply(ys, function(y){!all(is.na(y$.value)) & !is.na(var(y$.value,na.rm = T))})
ydata = do.call(rbind,ys)
colnames(ydata)[1] = ".id"
fit = fpca.sc(ydata=na.omit(ydata[ydata$.id %in% names(table(ydata$.id)[1:10]),]),
              var = TRUE, simul = FALSE)
```

```{r}
library(fdapace)
library(fastDummies)
vecs = apply(sleep,1,function(r){
  # Pull info
  id = as.character(r["Subject.ID"])
  start = ymd_hms(as.character(r["Night.Start"]))
  end = ymd_hms(as.character(r["Night.End"]))
  # Find CGM file
  df = read.csv(cgms[id])
  df$timestamp = round_date(ymd_hms(df$timestamp),"5 minutes")
  # Get the time between sensors
  interval = table(round(diff(df$timestamp)/60))
  interval = names(interval)[which.max(interval)]
  # Put glucose values on time grid, pull glucose values
  times = round_date(seq(start,end,by = paste(interval,"min")),"5 minutes")
  glucs = data.frame(cbind(times,df$sensorglucose[match(times,df$timestamp)]))
  glucs = glucs[,2]
  # Grid sequence for functional data
  arg = seq(0,length(glucs))
  return(list(glucs,arg))
})
ylist = sapply(vecs,"[[",1)
tlist = sapply(vecs,"[[",2)
keep = sapply(ylist, function(g){!all(is.na(g)) & !is.na(var(g,na.rm = T))})
X = sleep %>% select(Subject.ID:Average.BMI.Z.score)
X = X[keep,]
X = dummy_cols(X[,"Combined.Group.Category"],remove_selected_columns=T,remove_first_dummy=T)
FPCAobj <- FPCA(Ly=ylist[keep], Lt=tlist[keep],list(plot = TRUE))
VCAM(Lt=tlist[1:10], Ly=ylist[1:10],X[1:10])
```

@@

-   What is the best way to pick the order of basis functions, etc.?
-   Currently I'm forcing the glucose measures to be on the interval \[0,1\], which means that the period for the Fourier basis is "nighttime." Since those with shorter sleep periods will have fewer observations, I think this means that

>>>>

```{r}
ys = apply(sleep,1,function(r){
  # Pull info
  id = as.character(r["Subject.ID"])
  start = ymd_hms(as.character(r["Night.Start"]))
  end = ymd_hms(as.character(r["Night.End"]))
  # Find CGM file
  df = read.csv(cgms[id])
  id = paste(as.character(r["Subject.ID"]),as.character(r["Day"]),sep = "_")
  df$timestamp = round_date(ymd_hms(df$timestamp),"5 minutes")
  # Get the time between sensors
  interval = table(round(diff(df$timestamp)/60))
  interval = names(interval)[which.max(interval)]
  # Put glucose values on time grid
  times = round_date(seq(start,end,by = paste(interval,"min")),"5 minutes")
  glucs = df$sensorglucose[match(times,df$timestamp)]
  y = data.frame(".obs"=rep(id,length(glucs)),
                 ".index"=seq(0,1,length.out=length(glucs)),
                 ".value" = glucs)
  return(y)
})
# Find those where all glucose values are missing
# keep = sapply(ys, function(y){!all(is.na(y$.value)) & !is.na(var(y$.value,na.rm = T))})
ydf = do.call(rbind,ys)
colnames(ydf)[1] = ".id"
```

```{r}
fit = fpca.sc(ydata = ydf[ydf$.id %in% names(table(ydf$.id))[1:5],],pve = 0.95)
```

@@

```{r}
# Create a functional data object

basisobj = create.fourier.basis(rangeval=c(0, 1))
plot(basisobj)
```

@@

-   What is the best way to pick the order of basis functions, etc.? Visually?
-   Currently I'm forcing the glucose measures to be on the interval \[0,1\], which means that the period for the Fourier basis is "nighttime." Since those with shorter sleep periods will have fewer observations, I think this means that
